Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-06 paper_jasa.tex
updating comments

%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
%\begin{thm}\label{thm:power}
% Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
%\end{thm}
% 
%{ \em proof}
%
%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
%\begin{eqnarray*}
%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
%
%Thus expected power in equation \ref{power_exp} be
%\begin{eqnarray*}
% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
%  &=& E(1-p_D)^{m-1}   \\
%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_D}(0) \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%
\begin{figure}[hbtp]
   \centering
       \scalebox{.65}{\includegraphics{power_expected.pdf}}
       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$ and $\sigma = 12$. \hh{what is $m$? - would it make sense to show a curves for varying values of $m$?}}
       \label{fig:power_expected}
\end{figure}

@@

@@

This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case 2. Data is simulated with two sample sizes ($n=100, 300$), two standard deviations of the error ($\sigma=5, 12$), a variety of slopes ($\beta$), as given in Table  \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$. For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups. \green{ I don't understand the next two sentences.} For some parameter combinations we have more than three replications. Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots.

The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid. To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data. 
An example lineup from this experiment is shown in Figure \ref{fig:test_continuous}.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11. \green{Check this} This lineup, we think, is a difficult one. To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters. For the smallest slopes the difficulty is 0 (hardest) and for the largest slopes the difficulty is 4 (easiest). For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups. Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0. The order that the lineups are shown to participants is randomized. \green{Is this a correct description of the procedure?} Each subject saw a substantial number of difficult lineups. The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants. It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. \green{perhaps should go into so much detail about this, or keep this for the conclusions section.}

>>>>

%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
%\begin{thm}\label{thm:power}
% Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
%\end{thm}
% 
%{ \em proof}
%
%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
%\begin{eqnarray*}
%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
%
%Thus expected power in equation \ref{power_exp} be
%\begin{eqnarray*}
% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
%  &=& E(1-p_D)^{m-1}   \\
%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_D}(0) \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%
\begin{figure}[hbtp]
   \centering
       \scalebox{.65}{\includegraphics{power_expected.pdf}}
       \caption{Expected power of visual test  and the power of UMP test for sample size $n =100$ and $\sigma = 12$. \hh{what is $m$? - would it make sense to show a curves for varying values of $m$?} \blue{$m$=20 here. When $m$ changes the type-I error also changes. Thus I can produce the difference between power curves for each $m$. Do you suggest that?}}
       \label{fig:power_expected}
\end{figure}

@@

\green{Discuss the UMP test \blue{(discussed right after equation \ref{power_exp}) }, and give the reference for it here \blue { I am looking for a good reference. Do you have any suggestion?}.}

@@

This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case 2. Data is simulated with two sample sizes ($n=100, 300$), two standard deviations of the error ($\sigma=5, 12$), a variety of slopes ($\beta$), as given in Table  \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$. For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups. \green{ I don't understand the next two sentences.} For some parameter combinations we have more than three replications \blue{(since we run an extension of this experiment with more lineups for some parameter combinations}. Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots \blue{(we can ignore this now as I see that blocking discussion in section \ref{sec:category} is removed.) }.

The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid. To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data. 
An example lineup from this experiment is shown in Figure \ref{fig:test_continuous}.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11. \green{Check this} This lineup, we think, is a difficult one. To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters. For the smallest slopes the difficulty is 0 (hardest) and for the largest slopes the difficulty is 4 (easiest). For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups. Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0. The order that the lineups are shown to participants is randomized.\blue{The order is as described not randomized after. The randomization is done picking first combination of sample size and sd as well as picking lineups for each combination of parameter} \green{Is this a correct description of the procedure?} Each subject saw a substantial number of difficult lineups. The intention was noble, to get a better handle on the power of the lineups for small slopes, but practically this made is very tiring for the participants. It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. \green{perhaps should go into so much detail about this, or keep this for the conclusions section.}
